---
title: "PA1_template.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The file is downloaded from the web and the data set is read into R

```{r}
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"activity.zip")
unzip("activity.zip",exdir = "activity")
activity <- read.csv("C:/Users/Sony/Documents/activity/activity.csv")
```

## Calculation of Steps Per day and Histogram of steps per day

### Number of steps taken each day 

```{r}
library(dplyr)
activity1 <- group_by(activity,date)
activity2 <- summarise(activity1, steps = sum(steps,na.rm = TRUE))
activity2 <- as.data.frame(activity2)
head(activity2)
```

### Histogram of number of steps taken each day

```{r}
hist(activity2$steps,xlab = "No. of Steps",ylab = "Frequency",main = "Histogram of Number of Steps Per Day")
```

### Mean and Median of total number of steps taken per day

```{r}
activity3 <- as.data.frame(summarise(activity1, steps_mean = mean(steps,na.rm = TRUE),steps_median = median(steps,na.rm = TRUE)))
tail(activity3)
mean_steps <- mean(activity2$steps)
median_steps <- median(activity2$steps)
```

The average steps taken per day is `r mean_steps` and the median of steps taken per day is `r median_steps`

## Average daily activity pattern

### Time series Plot

The below graph shows the Time series plot of Average Number of steps taken for every 5 minute interval

```{r}
activity4 <- group_by(activity,interval)
activity4 <- as.data.frame(summarise(activity4,steps = mean(steps,na.rm = TRUE)))
plot(activity4$interval,activity4$steps,type = "l",xlab = "5 Min Interval",ylab = "Average No. of steps",main = "Time series Plot")
```


### 5 Min interval with maximum number of steps

```{r}
interval <- activity4[activity4$steps == max(activity4$steps),]
```

The Interval `r interval[1,1]` have the maximum number of steps i.e. `r interval[1,2]`

## Imputing Missing Values

### Number of Missing values

```{r}
na_value <- sum(is.na(activity$steps))
```

The Number of Missing values in the steps column is `r na_value`

### Strategy to fill the missing data and creating a new data set with missing values filled in

```{r}
activity4 <- group_by(activity,interval)
activity4 <- as.data.frame(summarise(activity4,steps = mean(steps,na.rm = TRUE)))
logi <- which(is.na(activity$steps))
activity5 <- activity
for(i in seq_along(logi)){
  interval1 <- activity5[logi[i],3]  
  activity5[logi[i],1] <- activity4[activity4$interval == interval1,2]
}
head(activity5)
```

### Histogram of new data set comparison of mean and median with previous data set after filling in the missing values

```{r}
library(dplyr)
activity6 <- group_by(activity5,date)
activity7 <- as.data.frame(summarise(activity6, steps = sum(steps)))
hist(activity7$steps,xlab = "No. of Steps",ylab = "Frequency",main = "Histogram of Number of Steps Per Day")
activity8 <- as.data.frame(summarise(activity6, steps_mean = mean(steps),steps_median = median(steps)))
head(activity8)
mean_steps1 <- mean(activity7$steps)
median_steps1 <- median(activity7$steps)
```

The Average number of steps taken per day after processing the missing values is `r mean_steps1` and the median number of steps taken per day after processing the missing values is `r median_steps1` . The value of average steps taken per day after processing the missing values has increased by `r mean_steps1 - mean_steps` and the median number of steps taken per day after processing the missing values has increased by `r median_steps1 - median_steps` .

## Difference in Activity pattern on Weekdays and Weekends

### Creating a new factor variable with two data sets = "weekday" and "Weekend" 

```{r}
weekdays1 <- c('Monday','Tuesday','Wednesday','Friday','Saturday')
activity5$Weekday <- factor(weekdays(as.POSIXct(activity5$date)) %in% weekdays1,levels = c(FALSE,TRUE),labels = c("weekend","weekday"))
```


### Creating a Panel Plot

```{r}
library(lattice)
act_weekday <- activity5[activity5$Weekday %in% "weekday",]
act_weekend <- activity5[activity5$Weekday %in% "weekend",]
act_weekday <- act_weekday[,1:3]
act_weekend <- act_weekend[,1:3]
act_weekday1 <- group_by(act_weekday,interval)
act_weekday1 <- as.data.frame(summarise(act_weekday1,steps = mean(steps)))
act_weekend1 <- group_by(act_weekend,interval)
act_weekend1 <- as.data.frame(summarise(act_weekend1,steps = mean(steps)))
act_weekday1$Weekday <- "weekday"
act_weekend1$Weekday <- "weekend"
final_act <- rbind(act_weekday1,act_weekend1)
xyplot(steps~interval | Weekday , data = final_act,type = "l",xlab = "Interval",ylab = "Number of steps",layout = c(1,2))
```

